Compositional Verification of Concurrent C Programs with Search Structure Templates

### 2023-05-02 concurrency/CONCUR 2023/paper.tex
Lennart's comments, second half

Our implementation of \lstinline{traverse} translates the functional implementation into more idiomatic C code. The lock-coupling pattern can be seen on lines 14 and 15, where \lstinline{traverse} acquires the next node's lock and then releases the current node's lock. The function stops when it reaches an empty leaf node (\lstinline{pn->p->t = NULL}), or when \lstinline{findNext} returns 0, indicating that the target key \lstinline{k} is in the current node. The \lstinline{traverse} function returns an integer flag that is 1 when we reach a leaf node and 0 when we find \lstinline{k} in an existing node; the behavior of operations that call \lstinline{traverse} can vary depending on this flag. For instance, an \lstinline{insert} operation may create a new node when the flag is 1 and modify an existing node when the flag is 0, while a \lstinline{lookup} operation may fail when the flag is 1 and return the value in the current node when the flag is 0.

@@

\subsubsection{Lock-coupling Proof}
\label{traverse_proof_lock}
%\wm{Does this organization work, or would it be better to prove each template as we introduce it?}
In this section, we show how to verify the templates presented above, by proving that each \lstinline{traverse} function meets its specification. The logically atomic specification for the lock-coupling template's \lstinline{traverse} function is shown in Figure \ref{fig:traverse_lock}. \wm{Note that this isn't the same as give-up, and that reflects the original paper. It's the cssOp/insert function that has the same spec in both.}
%\than{Try to avoid using specific name of $\mathrm{tree\_rep}$ for public condition, I suggested to use $\mathrm{Node}$ for generic purpose.}

@@

The give-up template's \texttt{traverse} specification, shown in Figure \ref{fig:traverse_giveup},  is simpler than that of the lock-coupling template. The local preconditions include any locks, and consist only of $\infp(\texttt{n})$, a footprint predicate asserting that node \texttt{n} is included in the search structure's abstract state $m$. As before, the \texttt{traverse} function will navigate the structure and return a node \texttt{n'} whose range includes \texttt{k}, as well as acquiring \texttt{n'}'s lock and returning its contents $\mathsf{R}(\texttt{n'})$. Also as before, the Boolean $\mathit{res}$ indicates whether \texttt{n'->t} is \texttt{NULL} or not.

@@

We instantiate the template with the binary search tree by defining and verifying the helper functions \texttt{findNext} (used in \lstinline{traverse}) and \texttt{insertOp} (used in \texttt{insert}). Their implementations are strictly sequential, their specifications are ordinary non-atomic Hoare triples, and their proofs do not require any concurrency reasoning.

>>>>

Our implementation of \lstinline{traverse} translates the functional implementation into idiomatic C code. The lock-coupling pattern can be seen on lines 14 and 15, where \lstinline{traverse} acquires the next node's lock and then releases the current node's lock. The function stops when it reaches an empty leaf node (\lstinline{pn->p->t = NULL}), or when \lstinline{findNext} returns 0, indicating that the target key \lstinline{k} is in the current node. The \lstinline{traverse} function returns an integer flag that is 1 when we reach a leaf node and 0 when we find \lstinline{k} in an existing node; the behavior of operations that call \lstinline{traverse} can vary depending on this flag. For instance, an \lstinline{insert} operation may create a new node when the flag is 1 and modify an existing node when the flag is 0, while a \lstinline{lookup} operation may fail when the flag is 1 and return the value in the current node when the flag is 0.

@@

\subsubsection{Lock-coupling Proof}
\label{traverse_proof_lock}
%\wm{Does this organization work, or would it be better to prove each template as we introduce it?}
We verify the template by proving that \lstinline{traverse} function meets a logically atomic specification, shown in Figure \ref{fig:traverse_lock}. \wm{Note that this isn't the same as give-up, and that reflects the original paper. It's the cssOp/insert function that has the same spec in both.}
%\than{Try to avoid using specific name of $\mathrm{tree\_rep}$ for public condition, I suggested to use $\mathrm{Node}$ for generic purpose.}

@@

The give-up template's \texttt{traverse} specification, shown in Figure \ref{fig:traverse_giveup},  is simpler than that of the lock-coupling template. The local precondition does not hold any locks, consisting only of $\infp(\texttt{n})$, a footprint predicate asserting that node \texttt{n} is included in the search structure's abstract state $m$. As before, the \texttt{traverse} function will navigate the structure and return a node \texttt{n'} whose range includes \texttt{k}, as well as acquiring \texttt{n'}'s lock and returning its contents $\mathsf{R}(\texttt{n'})$. Also as before, the Boolean $\mathit{res}$ indicates whether \texttt{n'->t} is \texttt{NULL} or not.

@@

We instantiate the templates with the binary search tree by defining and verifying the helper functions \texttt{findNext} (used in \lstinline{traverse}) and \texttt{insertOp} (used in \texttt{insert}). Their implementations are strictly sequential, their specifications are ordinary non-atomic Hoare triples, and their proofs do not require any concurrency reasoning.
